---
title: "Power Analysis"
author: "Vic Quennessen"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# load libraries
library(ggplot2)
library(viridis)
library(tidyverse)
library(data.table)
library(raster)
library(igraph)
library(rgeos)
library(gridExtra)
```

## Power analysis

Question 1: How many hatchlings should be sampled from a nest to robustly estimate the number of males that contributed to it?

### assuming random fertilization
```{r, fig.height = 2.25, warning = FALSE}
source('code/eggs_to_sample.R')

eggs_to_sample(n_eggs = 100, max_hatchlings = 96, max_males = 7, 
               breeding = 'random', n_sims = 1000)
```
\newpage

### assuming exponential decay in fertilization (1/2, 1/4, 1/8, etc.)
```{r, fig.height = 4}
source('code/eggs_to_sample.R')

eggs_to_sample(n_eggs = 100, 
               max_hatchlings = 96, 
               max_males = 7, 
               breeding = 'exponential', 
               n_sims = 1000)
```

\newpage

### assuming one dominant sire
```{r, fig.height = 4}
source('code/eggs_to_sample.R')

eggs_to_sample(n_eggs = 100, 
               max_hatchlings = 96, 
               max_males = 7, 
               breeding = 'dominate', 
               n_sims = 1000)
```

\newpage

Question 2: How many females and nests should be sampled to get a robust estimate of the number of breeding males, and therefore the breeding sex ratio?

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# biological parameters
max_n_males <- 5      # maximum number of males each female can mate with
max_n_females <- 10   # maximum number of females each male can mate with
minF  <- 10           # minimum number of females sampled
minM  <- 10           # minimum number of males in one breeding year
maxM  <- 60           # maximum number of males in one breeding year
trueF <- 60           # maximum number of females in one breeding year
Mprob <- rep(1/max_n_males, times = max_n_males)
# equal probability of mating with 1-5 males
mating_mode <- 'polyamory'
# options are 'polygyny' (males mate with multiple females but females mate with
# a single male) and 'polyamory' (both males and females have multiple partners)

# model parameters
nsims <- 100000

# dimensions
sampM <- seq(from = minM, to = maxM, by = 5)
nsampM <- length(sampM)
sampF <- seq(from = minF, to = trueF, by = 5)
nsampF <- length(sampF)

# pre-allocate data frame for results
DF <- data.frame(Number_of_males = rep(sampM, each = nsampF),
                 Sample_size = rep(sampF, times = nsampM),
                 Percent_identified = rep(NA, times = nsampF*nsampM),
                 Est_BSR = rep(NA, times = nsampF*nsampM))
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# for each number of potential males
for (m in 1:length(sampM)) {
  
  # pre-allocate vector of correct number of males identified
  # and breeding sex ratio
  percent <- matrix(NA, nrow = nsims, ncol = nsampF)
  BSR_estimate <- matrix(NA, nrow = nsims, ncol = nsampF)
  
  for (sim in 1:nsims) {
    
    # build pool of potential males to mate with
    Mpool <- rep(1:sampM[m], times = max_n_females)
    
    # simulate breeding - create a matrix of males mated with
    if (mating_mode == 'polygyny') {
      
      mates <- matrix(sample(Mpool, size = trueF, replace = FALSE), ncol = 1)
      
    } else if (mating_mode == 'polyamory') {
      
      # how many males does each female mate with
      n_males <- sample(1:max_n_males, size = trueF, prob = Mprob, 
                        replace = TRUE)
      
      # initialize mates matrix
      mates <- matrix(NA, nrow = trueF, ncol = max_n_males)
      
      # for each female, pull appropriate number of males from mating pool
      for (i in 1:trueF) {
        j <- n_males[i]
        mates[i, ] <- c(sample(Mpool, size = j), rep(NA, max_n_males - j))
      }
      
    }
    
    # identify total number of contributing males (correct value)
    trueM <- length(unique(as.vector(mates))) - 1
    
    # simulate sampling
    
    ## assuming perfect information
    
    # for each sample size of females, from 10 to total number of females 
    # nesting
    for (f in 1:length(sampF)) {
      
      # sample females to get represented males
      sampled_males <- length(unique(as.vector(na.omit(mates[sample(1:trueF, 
                                                                    size = sampF[f]), 
                                                             ]))))
      
      # determine percent of breeding males identified
      percent[sim, f] <- (sampled_males / trueM)
      
      # determine estimated breeding sex ratio
      BSR_estimate[sim, f] <- (sampled_males / trueF)
      
    }
    
    ## assuming imperfect information
    
  }
  
  # add proportion of times the correct number of males was IDd by number of 
  # males and sample size
  
  # take mean of correct columns (sample sizes) to get proportion correct
  percent_identified <- colMeans(percent, na.rm = TRUE)
  
  # take mean of correct columns (sample sizes) to get proportion correct
  avg_BSR_estimate <- colMeans(BSR_estimate, na.rm = TRUE)
  
  # calculate appropriate indices for DF
  index_start <- (m - 1)*nsampF + 1
  index_end <- index_start + nsampF - 1
  
  # store proportion correct values in DF
  DF$Percent_identified[index_start:index_end] <- percent_identified
  
  # store average BSM in DF
  DF$Est_BSR[index_start:index_end] <- avg_BSR_estimate
  
}
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# adjust sample_size to be proportional
DF$Prop_sampled <- DF$Sample_size / trueF

# add column for proportion correct > 80% TRUE or FALSE
DF$PC <- DF$Percent_identified >= 0.80

# convert data.frame to data.table and reshape to wide
DF2 <- dcast(setDT(DF), Number_of_males ~ Prop_sampled, value.var = "PC")

# reverse order of rows to match raster order
# remove first column
# convert to matrix and then to raster
r <- raster(as.matrix(DF2[ , .SD[.N:1, -1]]),
            xmn = 0, xmx = ncol(DF2) - 1, ymn = 0, ymx = ncol(DF2) - 1)

# detect clumps of connected cells of the value TRUE
# convert raster to polygons
# dissolve polygons into multi-polygons
polys <- rasterToPolygons(clump(r), dissolve = TRUE)

# grab coordinates of individual polygons and convert to a data.table
# use idcol = TRUE to enable grouping of paths when plotting
d_poly <- rbindlist(lapply(polys@polygons,
                           function(x) as.data.table(x@Polygons[[1]]@coords)),
                    idcol = TRUE)

# plot an outline around each 'patch of significant values' using geom_path
ggplot(data = DF, aes(x = Prop_sampled,
                      y = Number_of_males)) +
  geom_tile(aes(fill = Percent_identified)) +
  scale_fill_viridis(begin = 0, end = 1, direction = -1, discrete = FALSE,
                     breaks = c(0.1, 0.5, 0.9), labels = c(0.1, 0.5, 0.9))  +
  ylab('Number of potentially breeding males') +
  xlab('Proportion of females sampled') +
  labs(fill = 'Breeding \n sex ratio \n correctly \n identified') +
  geom_path(data = d_poly, aes(x = 0.25*x / 3 + 0.75/6, 
                               y = 5*y + 7.5, 
                               group = .id),
            size = 1, color = "brown1")

# save results to image file
ggsave(plot = last_plot(), filename = 'figure2.png',
       path = 'C:/Users/Vic/Box Sync/Quennessen_Thesis/presentations/
       NMFS Scholars Symposium 2021/',
       width = 4, height = 3)
```

```{r}
# figure 3 - breeding sex ratio heatmap
ggplot(data = DF, aes(x = Prop_sampled,
                      y = Number_of_males)) +
  geom_tile(aes(fill = Est_BSR)) +
  scale_fill_viridis(begin = 0, end = 1, direction = -1, discrete = FALSE,
                     breaks = c(0.2, 0.4, 0.6), labels = c(0.2, 0.4, 0.6))  +
  ylab('Number of potentially breeding males') +
  xlab('Proportion of females sampled') +
  labs(fill = 'Estimated \n breeding \n sex ratio')

# save results to image file
ggsave(plot = last_plot(), filename = 'figure3.png',
       path = 'C:/Users/Vic/Box Sync/Quennessen_Thesis/presentations/
       NMFS Scholars Symposium 2021/',
       width = 4, height = 3)
```

